---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---
Daily Activity Monitoring Analysis
-----------------------------------

* Dataset :   
    [Active Monitorng Data](https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip)

* The variables included in this dataset are:

    + steps: Number of steps taking in a 5-minute interval (missing values are coded as NA)
    + date: The date on which the measurement was taken in YYYY-MM-DD format
    + interval: Identifier for the 5-minute interval in which measurement was taken  
    
* The dataset is stored in a comma-separated-value (CSV) file and there are a total of 17,568 observations in this dataset.

```{r env setup, echo=FALSE}
setwd("C:/Users/a448166/Desktop/R-Course/Assignments/Project5")
```

## Data Processing
```{r data processing}
zipfilename <- "repdata_data_activity.zip"
file1 <- "activity.csv"

if (!file.exists(file1)){
  unzip(zipfilename) 
}

activity_data <- read.csv("activity.csv")

activity_data$date <- as.Date(activity_data$date)
activity_data$interval <- sprintf("%04d", activity_data$interval)
```
## Histogram of the total number of steps taken each day

### Calculate the total number of steps taken per day without missing values

```{r}
total_steps <- tapply(activity_data$steps,activity_data$date,sum)
```

```{r histogram}
library(ggplot2)
library(scales)
library(timeDate)
qplot(total_steps, geom = "histogram", hjust = 0.5 , binwidth = 1000, fill = ..count.., xlab = "Total steps per day", ylab = "Number of Days", main = "Total steps per day Frequency")
```

## Mean and median number of steps taken each day

```{r mean_median}
mean_steps <- mean(total_steps, na.rm = TRUE)
mean_steps
median_steps <- median(total_steps, na.rm = TRUE)
median_steps
```

## Time series plot of the average number of steps taken
```{r Time series plot}
split_data <- split(activity_data,activity_data$interval)
avg_steps <- sapply(split_data,function(par1) mean(par1$steps, na.rm = TRUE))
activity_data$avg_steps <- avg_steps

activity_data$interval_time <- strptime(activity_data$interval, format = "%H%M")

plot(activity_data$interval_time[1:288], avg_steps, type = "l", col = "red", lwd=3, main = "Average daily activity pattern", xlab = "5-minute Interval", ylab = "Average Number of steps")
```

## The 5-minute interval that, on average, contains the maximum number of steps

```{r maximum}
max_steps_ind <- which(avg_steps == max(avg_steps))
max_steps <- activity_data$interval[max_steps_ind]
max_steps
```

## Code to describe and show a strategy for imputing missing data

```{r impute missing data}
count_missing_data <- 0 

for (i in seq(nrow(activity_data))) {
  if (is.na(activity_data[i,"steps"])) {
    activity_data$steps_missingdata[i] <- activity_data[i,"avg_steps"]
    count_missing_data <- count_missing_data + 1
  }
  else {
    activity_data$steps_missingdata[i] <- activity_data[i,"steps"]
  }
}

count_missing_data
```

## Histogram of the total number of steps taken each day after missing values are imputed

### Calculate the total number of steps taken per day with missing values

```{r}
total_steps <- tapply(activity_data$steps_missingdata,activity_data$date,sum)
```

```{r histogram2}
qplot(total_steps, geom = "histogram", hjust = 0.5 , binwidth = 1000, fill = ..count.., xlab = "Total steps per day", ylab = "Number of Days", main = "Total steps per day with missing data")
```

## Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends  

### Mean and median number of steps taken each day

```{r mean_median_missing}
mean_steps <- mean(total_steps)
mean_steps
median_steps <- median(total_steps)
median_steps
```

### Distinguish between weekday and weekend

```{r weekday_weekend}
weekday <- isWeekday(activity_data$date)
weekend <- isWeekend(activity_data$date)

activity_data$day[weekday] <- "weekday"
activity_data$day[weekend] <- "weekend"

activity_data$day <- as.factor(activity_data$day)

split_data <- split(activity_data,activity_data$day)
weekday_weekend <- lapply(split_data,function(par1) split(par1,par1$interval))

weekday_data <- weekday_weekend[["weekday"]]
weekend_data <- weekday_weekend[["weekend"]]

weekday_avg_steps <- sapply(weekday_data,function(par1) mean(par1$steps, na.rm = TRUE))
weekend_avg_steps <- sapply(weekend_data,function(par1) mean(par1$steps, na.rm = TRUE))

weekday_plot_data <- data.frame(interval = activity_data$interval_time[1:288], steps = weekday_avg_steps, day = "weekday")
weekend_plot_data <- data.frame(interval = activity_data$interval_time[1:288], steps = weekend_avg_steps, day = "weekend")
plot_data <- rbind(weekday_plot_data, weekend_plot_data)
```

### Panel plot

```{r panel plot}
p <- ggplot(plot_data, aes(interval, steps))
p + geom_line(aes(colour = "red"), size = .7) + labs(x = "Interval", y="Number of steps", title = "Average daily activity pattern -weekday vs weekend")+scale_x_datetime(labels= date_format("%H:%M"))+theme (plot.title = element_text (hjust = 0.5))+facet_grid(day~.)
```
